---
title: "MTN Telco Custormer Rention analysis"
output:
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: sentence
---

## 1. Defining the Question

### a) Specifying the Data Analysis Question

The management would like to get your assistance in understanding the current subscribed customers.
Your recommendations informed by your analysis will help them make informed decisions on effective customer retention programs.

### b) Defining the Metric for Success

The solution will have satisfied the research question when you can identify the factors that contribute to customer churn.

### c) Understanding the context

MTN Telecom offers mobile and internet services to its customers.
These services include phone, multiple lines, internet, online security, online backup, device protection, tech support, and streaming TV and movies.

You have been provided with the current customer data.
Since you will be working towards a descriptive report than a predictive one, you decide to think critically of the kind of questions that would help you craft customer retention programs.
You then later use the given data set to answer your questions but before you start, you reading, explore, clean and visualise your dataset.

### d) Recording the Experimental Design

The steps to be taken include: 1.
Load dataset and preview its summarized information to get a feel of what you will be working with.
2.
Carry out data cleaning.
3.
Carry out data analysis.
4.
Interpret results.
5.
Provide recommendations based on results of analysis.
6.
Challenge your solution.

### e) Data Relevance

For now, the data we have contains churn data which will be critical for our research specific analysis.

## 2. Data Cleaning & Preparation

```{r}
#loading libraries
library(tidyverse)

```

```{r}
#loading data
df_mtn <- read_csv("https://bit.ly/2ZlpzjF")
df_mtn

```

```{r}
#data dimension
dim(df_mtn)

```

This data set has 7,050 records and 21 variables.

```{r}
#remove duplicates
df_mtn <- distinct(df_mtn)
dim(df_mtn)

```

7 duplicates values were found in the dataset which were dropped.

```{r}
#drop missing value
df_mtn <- na.omit(df_mtn)
dim(df_mtn)
```

40 missing values were dropped, which isn't a big loss.

```{r}
#properly formatting columns to lowercase
names(df_mtn) <- tolower(names(df_mtn))
names(df_mtn)

```

```{r}
#previewing the number of unique values
sample_n(unique(df_mtn), 7)
```

```{r}
#checking the datatypes
str(df_mtn)
```

```{r}
# check unique seniors
distinct_senior = unique(df_mtn$seniorcitizen)
distinct_senior
```

```{r}
# check unique seniors
distinct_payment = unique(df_mtn$paymentmethod)
distinct_payment
```

We can see that some variables aren't quite right when it comes to their unique values.
For example: \* 'seniorcitizen' should be a boolean object (Yes/No), not a float binary (0.0/0.1) \* 'paymentmethod' has values with spelling errors such as "Mailed checkkk" and "Electronic chekk", which created duplicates

```{r}
# resolve unique seniors
df_mtn$seniorcitizen <- replace(df_mtn$seniorcitizen,df_mtn$seniorcitizen ==0,"No")
df_mtn$seniorcitizen <- replace(df_mtn$seniorcitizen,df_mtn$seniorcitizen ==1,"Yes")
unique(df_mtn$seniorcitizen)
```

```{r}
# resolve payment methods
df_mtn$paymentmethod <- replace(df_mtn$paymentmethod,df_mtn$paymentmethod =="Mailed checkkk","Mailed check")
df_mtn$paymentmethod <- replace(df_mtn$paymentmethod,df_mtn$paymentmethod =="Electronic checkk","Electronic check")
unique(df_mtn$paymentmethod)
```

```{r}
# do a box plot on outliers
boxplot(select(df_mtn, tenure, monthlycharges, totalcharges))
```

We can see that there are several outliers in the 'tenure', 'monthly_charges', and 'total_charges' variables.
We can remove them using the IQR method.

```{r}
#find Q1, Q3, and interquartile range for values in column tenure
quartiles <- quantile(df_mtn$tenure, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(df_mtn$tenure)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

#only keep rows in dataframe that have values within 1.5*IQR of Q1 and Q3
data_no_outlier <- subset(df_mtn, df_mtn$tenure > Lower & df_mtn$tenure < Upper)

#view row and column count of new data frame
dim(data_no_outlier) 


```

## 3. Data Analysis

### Research-specific Analysis

What percentage of customers from our dataset churned?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn) %>% 
    mutate(percent = n/sum(n))
```

26.6% Churned

What is the gender role in churn?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn, gender) %>% 
    mutate(percent = n/sum(n))
```

Female and male have 13.36% and 13.23% respectively, indicating gender doesn't influence churn

Does senior citizen influence churn?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn, seniorcitizen) %>% 
    mutate(percent = n/sum(n))
```

Senior citizens seems to have a low 6.7% churn compared to non-senior citizens with 19.83%

Does having a partner influence churn?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn, partner) %>% 
    mutate(percent = n/sum(n))
```

Customers with partners have lower churn 9.5% compared to those without 17.07%

Does having a dependents influence churn?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn, dependents) %>% 
    mutate(percent = n/sum(n))
```

Customers with dependents have lower churn 4.6% compared to those without 21.94%

Does having a phone service influence churn?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn, phoneservice) %>% 
    mutate(percent = n/sum(n))
```

Customers with phone service have a high churn 24.17% compared to those without 2.4%

Does having multiple lines influence churn?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn, multiplelines) %>% 
    mutate(percent = n/sum(n))
```

There is no significant difference for the customers who have multiple lines while those customers without no phone service their churn is lower 2.4%.

Does having internet service influence churn?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn, internetservice) %>% 
    mutate(percent = n/sum(n))
```

Fiber optic customers have the highest churn rate of 18.43%

Does having online security influence churn?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn, onlinesecurity) %>% 
    mutate(percent = n/sum(n))
```

Customers without online security have the highest churn rate of 20.8%

Does having online backup influence churn?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn, onlinebackup) %>% 
    mutate(percent = n/sum(n))
```

Customers with online backup have the highest churn rate of 17.5%

Does having device protection influence churn?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn, deviceprotection) %>% 
    mutate(percent = n/sum(n))
```

Customers without device protection have the highest churn rate of 17.22%

Does having tech support influence churn?

```{r}
# count then perform percentage

df_mtn %>%
  count(churn, techsupport) %>% 
    mutate(percent = n/sum(n))
```

Customers that don't get tech support have the highest churn rate 20.59%

Does monthly in respect to payment method influence churn?

```{r}
# group by then perform summary

df_mtn %>%
  group_by(churn, paymentmethod) %>% 
    summarise(monthlycharges = sum(monthlycharges)) %>% arrange(desc(monthlycharges))
```

Electronic check payment method seems to have the highest churn followed by Bank transfer(automatic), Credit card (automatic) and Mailed check.

Does total charges in resepect to payment method influence churn?

```{r}
# count then perform percentage

df_mtn %>%
  group_by(churn, paymentmethod) %>% 
    summarise(totalcharges = sum(totalcharges)) %>% arrange(desc(totalcharges))
```

Electronic check and Bank transfer (automatic) payment methods even though they have the highest billing charges they top the payment methods that was used by the churned customers.

## 4. Summary of Findings

Based on the results of the analysis, the following conclusions were arrived at:

-    There is no significant difference in churn rate between male and female subscribers.
    So this is not an area management needs to worry about.

-   Majority of the customers are not senior citizens so this makes this dataset biased and hard to identify whether being a senior citizen affects churn rate.

-   Not having a partner increases the likelihood of churning.

-   Not having dependents increases the likelihood of churning.

-   Having Fiber optic as internet service increases the likelihood of churning.

-   Having phone service increases the likelihood of churning.

-   Having single or multiple line increases the likelihood of churning.

-   Having online backup increases the likelihood of churning.

-   Not having online security increases the likelihood of churning.

-   Having Electronic check and Bank transfer (automatic) as payment methods increases the likelihood of churning.

## 5. Recommendations

In order to create an effective customer retention program, management should take the following measures:

-   Focus more on meeting the needs of non-senior citizens.

-   Focus more on having customers that have partners and/or dependents since these people are less likely to churn.
    Alternatively, management can come up with services specifically designed for customers without partners and/or dependents.
    This would require additional research.

-   Focus on providing more online security services to customers.
    *Encourage customers to user Credit card as payment method. More research is required to understand why Electronic check has the highest churn rate.*

-   Focus on marketing more on internet services to address the significant number of customers with no internet service.

-   Focus on providing device protection services to customers

-   Focus on improving customer service in respect to tech support.

##6.
Challenging your Solution

### a). Did we have the right data? Do we need other data to answer our question?

s far as I can tell, we had the right data.
However, more data is still needed, particularly those with more customers who churned so we can have a better understanding of why they might have churned.

### b). Did we have the right question?

Yes, we did.
